---
title: "Introduction to R"
tutorial: 
  id: "introR"
output: learnr::tutorial
runtime: shiny_prerendered
description: Check out your R skills for our introduction to data analytics (introduction to statistics!) course. 
---

```{r setup, include=FALSE}
library(learnr)
library(learnSTATS)
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction to R Class Assignment

In this example, you will practice your *R* skills! You will map key concepts of variable types to objects types and begin interacting in a console as part of this tutorial. The learning objectives include:

- Understand logical operators
- Work with variables, vectors, matrices, and dataframes
- Write your own functions
- Load libraries and files 
  
## Introduction R Video Part 1

The following videos are provided as a lecture for the class material. You can view the lecture notes within *R* using `vignette("Introduction-to-R", "learnSTATS")`. You can skip these pages if you are in class to go on to the assignment. 

<iframe width="560" height="315" src="https://www.youtube.com/embed/oV0SBUbmnmg" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Introduction R Video Part 2

<iframe width="560" height="315" src="https://www.youtube.com/embed/W_xmhVzF12s" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Introduction R Video Part 3

<iframe width="560" height="315" src="https://www.youtube.com/embed/NuDxrW4jsgQ" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Exercises

In this next section, you will answer questions using the *R* code blocks provided. Be sure to use the `solution` option to see the answer if you need it!

Please enter your name for submission. If you do not need to submit, just type anything you'd like in this box. 

```{r details}
question_text(
  "Student Name:",
  answer("Your Name", correct = TRUE),
  incorrect = "Thanks!",
  try_again_button = "Modify your answer",
  allow_retry = TRUE
)
```

## Logical Operators 

Use logical operations to get R to agree that "two plus two equals 5" is FALSE.

```{r q1, exercise = TRUE}

```

```{r q1-solution}
2 + 2 == 5
```

Use logical operations to test whether 8 ^ 13 is less than 15 ^ 9.

```{r q2, exercise = TRUE}

```

```{r q2-solution}
8 ^ 13 < 15 ^ 19
```

## Simple Variables

Create a variable called `potato` whose value corresponds to the number of `potatoes` you've eaten in the last week. Then print out the value of `potato`. 

```{r q3, exercise = TRUE}

```

```{r q3-solution}
#there are a lot of right answers!
potato <- 15
```

Calculate the square root of `potato` using the `sqrt()` function. Print out the value of potato again to verify that the value of potato hasn't changed. Note that `potato` has been set to 15 for this question. 

```{r q4-setup}
potato <- 15
```

```{r q4, exercise = TRUE}

```

```{r q4-solution}
sqrt(potato)
potato
```

Reassign the value of `potato` to `potato * 2`. Print out the new value of `potato` to verify that it has changed.

```{r q5-setup}
potato <- 15
```

```{r q5, exercise = TRUE}

```

```{r q5-solution}
potato <- potato * 2
potato
```

Create three new variables:
  
  - Create a string (character variable), and name that variable `charVar`. 
  - Create a logical variable, and name that variable `logicalVar`.
  - Create a variable that contains a missing value `NA`, and name that variable `missingVar`. 
  - Use `class()` to verify that each variable is the right type. 

```{r q6, exercise = TRUE}

```

```{r q6-solution}
charVar <- "cheese"
logicalVar <- TRUE
missingVar <- NA
class(charVar)
class(logicalVar)
class(missingVar)
```

## Vectors

Create a numeric vector with three elements using c().

```{r q7, exercise = TRUE}

```

```{r q7-solution}
#lots of right answers!
c(1,2,3)
```

Create a character vector with three elements using c().

```{r q8, exercise = TRUE}

```

```{r q8-solution}
#lots of right answers!
c("this", "is", "characters")
```

Create a numeric vector called `age` whose elements contain the ages of three people you know. Add the names of these people using `names(age)`. 

```{r q9, exercise = TRUE}

```

```{r q9-solution}
#lots of right answers!
age <- c(34, 38, 27)
names(age) <- c("scott", "abby", "katy")
```

Use "indexing by number" to get R to print out the first element of `Orange$circumference`.

```{r q10, exercise = TRUE}
data("Orange")

```

```{r q10-solution}
data("Orange")
Orange$circumference[1]
```

Use logical indexing to return all the circumferences of Orange trees greater than 100.

```{r q11, exercise = TRUE}
Orange$circumference[]
```

```{r q11-solution}
Orange$circumference[Orange$circumference > 100]
```

## Dataframes

Load the `airquality` dataset. Use the $ method to print out the `Wind` variable in `airquality.`

```{r q13, exercise = TRUE}

```

```{r q13-solution}
data("airquality")
airquality$Wind
```

Print out only the first 10 cases (rows) in the `airquality` dataset. Hint: typing c(1,2,3,4,5,6,7,8,9,10) is tedious. R allows you to use 1:10 as a shorthand method!

```{r q16, exercise = TRUE}

```

```{r q16-solution}
airquality[1:10 , ]
```

Use logical indexing to print out all days in `airquality` where the `Ozone` level was higher than 20. 

```{r q17, exercise = TRUE}

```

```{r q17-solution}
airquality[airquality$Ozone > 20 , ]
```

Use subset() to do the same thing. Notice the difference in the output.

```{r q18, exercise = TRUE}

```

```{r q18-solution}
subset(airquality, Ozone > 20)
```

Use the `length()` function to determine the number of observations in the `airquality` dataframe.

```{r q20, exercise = TRUE}

```

```{r q20-solution}
#remember that number of rows needs to be on a vector 
length(airquality$Ozone) 
```

## Functions 

Make a simple function to calculate x + 6.

```{r q24, exercise = TRUE}

```

```{r q24-solution}
addup <- function(x) { x + 6 }
```

## Packages

Load the `utils` library.

```{r q29, exercise = TRUE}

```

```{r q29-solution}
library(utils)
```

## Submit

On this page, you will create the submission for your instructor (if necessary). Please copy this report and submit using a Word document or paste into the text window of your submission page. Click "Generate Submission" to get your work! 

```{r context="server"}
encoder_logic()
```

```{r encode, echo=FALSE}
encoder_ui()
```
