---
title: "Correlation"
tutorial:
  id: "correlation"
output: learnr::tutorial
runtime: shiny_prerendered
description: Let's now move on from basic statistics to building our first models using the mean and standard error. 
---

```{r setup, include=FALSE, mesage = F}
library(learnr)
library(learnSTATS)
library(faux)

##original data
lab7_original <- read.csv(paste("07_lab_correlation/07_data", current_data, sep = ""))
summary(lab7_original)

lab7_update <- sim_df(lab7_original, #data frame
                      how_many, #how many of each group
                      between = "gender",
                      seed = this_year_seed) 

##add a bit more noise
lab7_update$temporality <- lab7_update$temporality * small_changes[1]
lab7_update$expectability <- round(lab7_update$expectability * small_changes[2], 0)
lab7_update$relativity <- lab7_update$relativity * small_changes[3]
lab7_update$positive <- lab7_update$positive * small_changes[4]



lab7_update <- messy(lab7_update, 
                     prop = .03,
                     3:6,
                     replace = NA)

knitr::opts_chunk$set(echo = FALSE)
```

## Introduction to More Statistics

In this example, we will begin to build our first statistical models. You will learn how to use `apply` family of functions to calculate our descriptive statistics. The learning objectives include:

- Know what a statistical model is and why we use them
- Measures of central tendency, dispersion, frequency distributions
- Know what the 'fit' of a model is and why it is important
- Distinguish models for samples and populations
  

https://youtu.be/LPlDgz6CNBE
https://youtu.be/zD5op6nHSpU

## More Statistics Video Part 1

The following videos are provided as a lecture for the class material. The lectures are provided here as part of the flow for the course. You can view the lecture notes within *R* using `vignette("More-Stats-Concepts", "learnSTATS")`. You can skip these pages if you are in class to go on to the assignment. 

<iframe width="560" height="315" src="https://www.youtube.com/embed/o5a7wQXtAe0" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## More Statistics Video Part 2

<iframe width="560" height="315" src="https://www.youtube.com/embed/mmRejGb3cYg" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Exercises

In this next section, you will answer questions using the *R* code blocks provided. Be sure to use the `solution` option to see the answer if you need it!

Please enter your name for submission. If you do not need to submit, just type anything you'd like in this box. 

```{r details}
question_text(
  "Student Name:",
  answer("Your Name", correct = TRUE),
  incorrect = "Thanks!",
  try_again_button = "Modify your answer",
  allow_retry = TRUE
)
```
